The Correctness of Algebraic File Synchronization

### 2015-11-26 2015/paper.tex
More proofs

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\wr\!\!\wr\,} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes

@@

% TODO delete:
\newcommand{\infer}[2]{\begin{array}{c}{#1}\\\hline{#2}\end{array}}
\newcommand{\inferr}[3]{\begin{array}{c}{#1}\\{#2}\\\hline{#3}\end{array}}
\newcommand{\inferrr}[4]{\begin{array}{c}{#1}\\{#2}\\{#3}\\\hline{#4}\end{array}}

@@

% TODO delete:
\newtheorem{myrul}{Rule}
\crefname{myrul}{rule}{rules}

\title{Correctness of Algebraic File Synchronization}

@@

% <<<<<<<< TODO Rewrite: no algebra
 about which we will reason
using our algebra.
We will aim to draw conclusions or judgements about commands and sequences of commands
based on the algebra we are aiming to construct.
For example, if we believe that two sequences of commands are equivalent as their effect
is the same, then we aim to be able to derive this from the axioms and inference rules of the
algebra.
As the algebra operates on commands only,
these judgements, by definition, need to hold regardless of the filesystem the commands are applied to.
Accordingly, we expect that the more information one encodes into the commands and the sequences,
the better predicitons we will be able to make using our algebra,
as then the commands and sequences will be more specific, and will select a smaller subset
of potential filesystems on which they can meaningfully operate.
% >>>>>>>>>> to here

@@

% TODO Proof of correctness of update detection - the generated order will work

@@

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in
the simple set $U$
and represent all orders generated by this algorithm.
\end{mydef}

It follows from the algorithm that:
\begin{mylem}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a $U'\in\orderset{U}$ where they are next to each other.
\end{mylem}

@@

\begin{mylem}\label{simple_is_max}
For any simple sequence $S\neq\cbrk$, $\forall T: S\eqext T \Rightarrow S\equiv T$;
that is, the domains of simple sequences cannot be extended further.
This is equivalent to the proposition that $\extset{S} = \{S'|S\equiv S'\}$.
\end{mylem}
\begin{proof}
TODO
% TODO
\end{proof}

@@

\begin{mydef}{Reconciliation}\label{def:reconciliation}
\begin{align*}
\reca = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: c\indep c'\\
\} &
\end{align*}

In other words, the sequence $\reca$ we apply to $\FS_B$ consists of all commands from $A\setminus B$
that are independent of every command in $B\setminus A$.
The order of commands in the sequence can be determined using the algorithm described in
\cref{ordering} as any subset of $A$ is also simple,
and $\recb$ can be obtained in a similar way by reversing $A$ and $B$.
\end{mydef}

We would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\FS_B\neq\fsbroken. \]
So that we can formalise statements needed to prove this,
let us introduce two relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

@@

\begin{mycor}
% An inference rule in the algebra
$B\eqext A \Rightarrow \worksc{A}{B}$
\end{mycor}

\begin{mycor}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksc{A}{A\cc S}$
\end{mycor}

\begin{mycor}
% An inference rule in the algebra
$\workssign$ can be chained:
if $A_1\in\seqset{A}$, then
$ \worksc{\seqset{A}}{\seqset{B}} \wedge \worksc{\seqset{C}}{A_1} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}} $
\end{mycor}

We will also prove the following lemmas.

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxy(n, V_Y)\in A$) and $B$ ($\czw(n, V_W)\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
$\cxy(n, V_Y)$ or $\czw(n, V_W)$ would not break, and consequently $A$ and $B$ could not work on the same filesystem.
\end{proof}

\begin{mylem}\label{r_indep_combine}
% An inference rule in the algebra
\[ \worksc{T}{\seqset{S}} \wedge \worksc{U}{\seqset{S}} \wedge T\indep U \Rightarrow \worksc{T\cc U}{\seqset{S}} \]
\end{mylem}
\begin{proof}
% TODO Proof
TODO
\end{proof}

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
$\cxy(n,V_Y)\indep\czw(m,V_W) \Leftrightarrow n\unrel m$ 
if $\cxy(n,V_Y)\neq\czw(m,V_W)$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
$\cxy(n,V_Y)\not\indep\czw(m,V_W) \wedge n\unrel m$ would lead to contradiction.
By definition, if the two commands are not independent, then one of the following holds:
\begin{itemize}
\item $\czw(m,V_W)\cc \cqr(o,V_R) \equiv\cbrk$
\item $\cqr(o,V_R)\cc \czw(m,V_W) \equiv\cbrk$
\item $\czw(m,V_W)\cc \cqr(o,V_R) \nequiv \cqr(o,V_R)\cc \czw(m,V_W)$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

Next, we prove that
$\cxy(n,V_Y)\indep\czw(m,V_W) \wedge n\not\unrel m$ would lead to contradiction.
As $n\not\unrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

@@

@@

\begin{mylem}\label{r_invmove}
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \\
\worksc{\seqset{A}}{B\cc \seqset{S}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{S}}
\end{gather*}
% TODO Proof
TODO.
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{mylem}

\begin{mylem}\label{indep_prefix_combine}
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{r_invmove} we know
$\worksc{B}{A^{-1}\cc \seqset{S}}$ and $\worksc{C}{A^{-1}\cc \seqset{S}}$,
and from \cref{r_indep_combine} we know
$\worksc{B\cc C}{A^{-1}\cc \seqset{S}}$. Applying the second line of \cref{r_invmove} we get
$\worksc{(A^{-1})^{-1}\cc B\cc C}{\seqset{S}}$, which proves our lemma as $(A^{-1})^{-1}\equiv A$.
\end{proof}

@@

\begin{myth}
If $A, B$ are simple, then $\worksc{B\cc \reca}{A,B}$
\end{myth}
where, to restate the definition,
{ % scope
\renewcommand{\themydef}{\ref{def:reconciliation}}
\begin{mydef}
\begin{align*}
\reca = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: c\indep c'\\
\}. &
\end{align*}
\end{mydef}
} % scope

@@

Let us consider therefore the marked command preceded by an unmarked command in $A$.
and let the marked command be $\cxy(n, V_Y)$, 
and the preceding unmarked command be $\czw(m, V_W)$.
As $A$ is a simple command and $\works{A}$, from 
\cref{ax_distantrel_breaks,ax_directchild_breaks,ax_directparent_breaks}
we know that these commands can only be on unrelated paths or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $A$,
and we show that the last two cases are impossible as they would lead to contradiction.

@@

\newcommand{\acb}{(A\cap B)}
\newcommand{\acbi}{\acb^{-1}}
\newcommand{\amb}{(A\setminus B)}
\newcommand{\bma}{(B\setminus A)}
We would like to prove that $\worksc{B\cc \reca}{A,B}$.
From the above we know that this is equivalent to
$\worksc{\acb\cc \bma\cc \reca}{A,B}$.
As we already know that $\reca\indep\bma$
and that $\worksc{\acb\cc \bma}{A,B}$
(as $\worksc{B}{A,B}$),
based on \cref{indep_prefix_combine}
we only need to prove that
$\worksc{\acb\cc \reca}{A,B}$.

Using \cref{r_invmove} we can rephrase this as
$\worksc{\reca}{\acbi\cc A,\acbi\cc B}$.

To restate some results above, we therefore already know the following things:
\begin{itemize}
\item $\amb$ and $\bma$ are simple as $A$ and $B$ are simple
\item $\amb \cap \bma = \emptyset$
\item $\reca$ is the largest subset of $\amb$ for which $\reca\indep\amb$
\item $\worksc{\amb, \bma}{\acbi\cc A,\acbi\cc B}$ 
based on \cref{r_invmove}
as $\worksc{A,B}{A,B}$
and $A\equiv\acb\cc\amb$, $B\equiv\acb\cc\bma$.
\end{itemize}

@@

% TODO $\reca$ is maximal

% TODO Talk about setting up an algebra / calculus

>>>>

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\,{\wr\wr}\,} % Independent commands, sequences
\newcommand{\unrel}{\indep} % Unrelated nodes
\newcommand{\nindep}{\,{\centernot{\wr\wr}}\,}
\newcommand{\nunrel}{\nindep}

@@

@@

\title{The Correctness of Algebraic File Synchronization}

@@

% <<<<<<<< TODO Rewrite: no algebra
TODO: Rewrite:
 about which we will reason
using our algebra.
We will aim to draw conclusions or judgements about commands and sequences of commands
based on the algebra we are aiming to construct.
For example, if we believe that two sequences of commands are equivalent as their effect
is the same, then we aim to be able to derive this from the axioms and inference rules of the
algebra.
As the algebra operates on commands only,
these judgements, by definition, need to hold regardless of the filesystem the commands are applied to.
Accordingly, we expect that the more information one encodes into the commands and the sequences,
the better predicitons we will be able to make using our algebra,
as then the commands and sequences will be more specific, and will select a smaller subset
of potential filesystems on which they can meaningfully operate.
% >>>>>>>>>> to here

@@

% TODO
TODO: Proof of correctness of update detection - the generated order will work

@@

\begin{mydef}{$\orderset{U}$.}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in
the simple set $U$
and represent all orders generated by this algorithm.
\end{mydef}

It follows from the algorithm that:
\begin{mycor}\label{lemma:neighbor}
If $U$ contains commands on both $n$ and $\parent(n)$, then
there is a $U'\in\orderset{U}$ where they are next to each other.
\end{mycor}

@@

\begin{mylem}\label{simple_is_max}
For any simple sequence $S\neq\cbrk$, $\forall T: S\eqext T \Rightarrow S\equiv T$;
that is, the domains of simple sequences cannot be extended further.
This is equivalent to the proposition that for simple sequences $\extset{S} = \{S'|S\equiv S'\}$.
\end{mylem}
\begin{proof}
TODO
% TODO
\end{proof}

@@

\begin{mydef}{Reconciliation.}\label{def:reconciliation}
\begin{align*}
\reca = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: c\indep c'\\
\} &
\end{align*}

In other words, the sequence $\reca$ we apply to $\FS_B$ consists of all commands from $A\setminus B$
that are independent of every command in $B\setminus A$.
The order of commands in the sequence can be determined using the algorithm described in
\cref{ordering} as any subset of $A$ is also simple.
In a similar way, $\recb$ can be obtained by reversing $A$ and $B$.
\end{mydef}

We would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\FS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce two relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

@@

\begin{mycor}
% An inference rule in the algebra
$B\eqext A \Rightarrow \worksc{A}{B}$.
\end{mycor}

\begin{mycor}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksc{A}{A\cc S}$, that is, if a sequence works, its prefix also works.
\end{mycor}

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\workssign$ can be chained:
if $A_1\in\seqset{A}$, then
$ \worksc{\seqset{A}}{\seqset{B}} \wedge \worksc{\seqset{C}}{A_1} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}}$.
\end{mycor}

We also prove the following lemmas.

\begin{mylem}\label{unrelated_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are unrelated:
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \Leftrightarrow n\unrel m \]
if $\cxy(n,V_Y)\neq\czw(m,V_W)$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
We proceed in two steps; first, we prove that
\[ \cxy(n,V_Y)\nindep\czw(m,V_W) \wedge n\unrel m \] would lead to contradiction.
By definition, if the two commands are not independent, then one of the following holds:
\begin{itemize}
\item $\czw(m,V_W)\cc \cqr(o,V_R) \equiv\cbrk$
\item $\cqr(o,V_R)\cc \czw(m,V_W) \equiv\cbrk$
\item $\czw(m,V_W)\cc \cqr(o,V_R) \nequiv \cqr(o,V_R)\cc \czw(m,V_W)$
\end{itemize}
However, from \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
if $n\unrel m$ then none of the above can be true.

Next, we prove that
\[ \cxy(n,V_Y)\indep\czw(m,V_W) \wedge n\nunrel m \] would lead to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\begin{align*}
\forall\FS: &\cxy(n,C_Y)\indep \czw(m,C_W) \\
&\wedge \cxy(n,C_Y)\FS\neq\fsbroken \\
&\wedge \czw(m,C_W)\FS\neq\fsbroken \\
\Rightarrow &\cxy(n,C_Y)\cc \czw(m,C_W)\FS\neq\fsbroken.
\end{align*}
\end{myax}
\begin{proof}
We name this proposition an axiom for the same reason as above,
but it is easy to provide an indirect proof based on our filesystem model.
Assume that $\cxy(n,C_Y)\cc \czw(m,C_W)\FS=\fsbroken$.
We know $\cxy(m,C_Y)\FS\neq\fsbroken$ so it must be applying 
$\czw(m,C_W)$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if there the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxy(m,C_Y)\FS)(m)\in\setvx{W}$ as based on \cref{unrelated_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxy(m,C_Y)$.
\end{proof}

This result can be extended to sequences:

\begin{mylem}\label{combine_independent_sequences}
The combination of independent sequences works wherever the original sequences work:
\[ S\indep T \Rightarrow \worksc{S\cc T}{S,T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\FS\neq\fsbroken$ and $T\FS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

From \cref{unrelated_is_independent,ax_separate_commute} we know that
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ preserving their original order,
but in any permutation between $S$ and $T$ is equivalent to $S\cc T$ on all filesystems.

Let the command that breaks $\FS$ in $T$ be $t$
so that $T=T_0\cc t\cc T_1$.
It is still true that $(T_0 \cc t)\FS\neq\fsbroken$,
and by definition $(S\cc T_0)\FS\neq\fsbroken$,
but $(S\cc T_0\cc t)\FS=\fsbroken$.
Also, from above we know that $S\cc T_0\equiv T_0\cc S$
and so $(T_0 \cc S)\FS\neq\fsbroken$.

If we denote the first command in $S$ with $s_1$,
this means that $(T_0 \cc s_1)\FS\neq\fsbroken$,
which we can combine with $(T_0 \cc t)\FS\neq\fsbroken$, $t\indep s_1$ and
\cref{combine_independent_commands}
(using $T_0\FS$ as the reference filesystem)
to arrive at $(T_0 \cc s_1\cc t)\FS\neq\fsbroken$.

We can repeat this step for $s_2$, the next command in $S$,
and from 
$(T_0 \cc s_1\cc t)\FS\neq\fsbroken$
and
$(T_0 \cc s_1\cc s_2)\FS\neq\fsbroken$
arrive at
$(T_0 \cc s_1\cc s_2\cc t)\FS\neq\fsbroken$.
This can be repeated until $S$ is exhausted and we get
$(T_0 \cc S\cc t)\FS\neq\fsbroken$, which is a contradiction.

\end{proof}

From this and \cref{workschained} it follows that:

\begin{mycor}\label{r_indep_combine}
% An inference rule in the algebra
The combination of independent sequences continues to work under the same conditions:
\[ \worksc{S}{\seqset{U}} \wedge \worksc{T}{\seqset{U}} \wedge S\indep T \Rightarrow \worksc{S\cc T}{\seqset{U}} \]
\end{mycor}

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxy(n, V_Y)\in A$) and $B$ ($\czw(n, V_W)\in B$)
then their input types must match ($X=Z$).
\end{mylem}
\begin{proof}
This is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
either $\cxy(n, V_Y)$ or $\czw(n, V_W)$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

\begin{mylem}\label{r_invmove}
A common prefix of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{S}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{S}} \\
\worksc{\seqset{A}}{B\cc \seqset{S}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{S}}
\end{gather*}
% TODO Proof
TODO.
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{mylem}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tail continues to work:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{r_invmove} we know
$\worksc{B}{A^{-1}\cc \seqset{S}}$ and $\worksc{C}{A^{-1}\cc \seqset{S}}$,
and from \cref{r_indep_combine} we know
$\worksc{B\cc C}{A^{-1}\cc \seqset{S}}$. Applying the second line of \cref{r_invmove} we get
$\worksc{(A^{-1})^{-1}\cc B\cc C}{\seqset{S}}$, which proves our lemma as $(A^{-1})^{-1}\equiv A$.
\end{proof}

@@

\begin{myth}
If $A, B$ are simple, then $\worksc{B\cc \reca}{A,B}$
\end{myth}
where, to restate \cref{def:reconciliation},
% { % scope
% \renewcommand{\themydef}{\ref{def:reconciliation}}
% \begin{mydef}
\begin{align*}
\reca = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: c\indep c'\\
\}. &
\end{align*}
% \end{mydef}
% } % scope

@@

Let us consider therefore the marked command preceded by an unmarked command in $A$,
and let the marked command be $\cxy(n, V_Y)$, 
and the preceding unmarked command be $\czw(m, V_W)$.
As $A$ is simple and $\works{A}$, from 
\cref{ax_distantrel_breaks,ax_directchild_breaks,ax_directparent_breaks}
we know that these commands can only be on unrelated paths or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $A$,
and we show that the last two cases are impossible as they would lead to contradiction.

@@

\newcommand{\acb}{(A\cap B)}
\newcommand{\acbi}{\acb^{-1}}
\newcommand{\amb}{(A\setminus B)}
\newcommand{\bma}{(B\setminus A)}
We would like to prove that $\worksc{B\cc \reca}{A,B}$.
From the above we know that this is equivalent to
\[ \worksc{\acb\cc \bma\cc \reca}{A,B}. \]
As we already know that $\reca\indep\bma$
and that $\worksc{\acb\cc \bma}{A,B}$
(as $\worksc{B}{A,B}$),
based on \cref{indep_prefix_combine}
we only need to prove that
\[ \worksc{\acb\cc \reca}{A,B}. \]
Using \cref{r_invmove} we can rephrase this as
\[ \worksc{\reca}{\acbi\cc A,\acbi\cc B}. \]

To restate some results above, we therefore already know the following things:
\begin{itemize}
\item $\amb$ and $\bma$ are simple as $A$ and $B$ are simple
\item $\amb \cap \bma = \emptyset$
\item $\reca$ is the largest subset of $\amb$ for which $\reca\indep\amb$
\item $\worksc{\amb, \bma}{\acbi\cc A,\acbi\cc B}$ 
based on \cref{r_invmove}
as $\worksc{A,B}{A,B}$,
and $A\equiv\acb\cc\amb$, and $B\equiv\acb\cc\bma$.
\end{itemize}

@@

% TODO
TODO: $\reca$ is maximal

% TODO
TODO: About setting up an algebra / calculus
